---
title: "Module 4 - Data Finding and Wrangling"
subtitle: "Gender and Nicotine Dependence"
author: Betelihem Getachew
date: "`r Sys.Date()`"
output:
  html_document:
    toc: FALSE
editor: 
  markdown: 
    wrap: 72
---

**Background**

The below data was retrieved from the Global Adult Tobacco Survey
website ([Dataset for African Region, Botswana, Botswana -
National](https://nccd.cdc.gov/GTSSDataSurveyResources/Ancillary/DataReports.aspx?CAID=2&Survey=4&WHORegion=3&Country=123&Site=27000)
). The data was collected in 2017 in Botswana among adults 15 years and
older. Both smokers and non smokers participated in the survey. a total
of 4643 participated in this survey, a subset of 463 participants are
used in this analysis. The below are unweighted calculations . The code
book(BOT_GATS_207...) is located in the data_analysis_exercise folder.

**Objective**

The objective of this analysis is to see if Gender predicts nicotine
dependence. Only smokers were included in this analysis. Heavy Smoking
Indeix(HSI) re:nicotine dependence was calculated as a score by adding
B01+B07 variables.The scores were categorized as low addicition if score
is between 0 and 2, and medium/high addiction score if between 3 and 6.

This uses MS Word as output format. [See
here](https://quarto.org/docs/output-formats/ms-word.html) for more
information. You can switch to other formats, like html or pdf. See [the
Quarto documentation](https://quarto.org/) for other formats.

**Data import and cleaning**

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(dplyr)
library(tidyr)
library(skimr)
library(haven)
library(memisc)
library(ggplot2)
library(labelled)
```

```{r}
Botsdata<-read_sav("data_analysis_exercise/Data/Botswana.sav")
```

```{r}
class(Botsdata)
```

#subsetting for variables of interest see bodebook. #A01 Gender, B01
smoking status, B04 age of smoking initiation, B07 smoking how soonafter
awaking up, B06A number of cigarettes smoked

```{r}
Bselect<- Botsdata[,c("A01","B01","B04","B07","B06A")]
```

#checking my variables names of interest are selected and their strucutre

```{r}
names(Bselect)
str(Bselect)
```
#labeling my variable names 
```{r}
my_labels <- c(A01 = "Gender",
               B01 = "Do you currenlty smoke?",
               B04 = "How old were you when you frist started smoking?",
               B07 = "How soon do you smoke after you wake up",
               B06A="How many cigarattes do you smoke each day?")
              
```

```{r}
print(my_labels)
```
#trying to set the variable name in the df but it didnt work
```{r}
Bselect<-set_variable_labels(Bselect,.labels = my_labels)
```


```{r}
summary(Bselect)
```

#creating a subset for smokers only based on B01 Do you currently smoke?
only interested in responses 1,2 only (for daily and less than daily)

```{r}
subset(Bselect, B01<=2)
```

#saving the above as your dataset

```{r}
Bselect2=subset(Bselect, B01<=2)
```

```{r}
summary(Bselect2)
```

#creting the HSI variable by adding B01 and B07,

```{r}
Bselect2$HSI<-Bselect2$B01 + Bselect2$B07
```

#check if your variable has been calculated/ added

```{r}
dim(Bselect2)
names(Bselect2)
str(Bselect2)
```

#check your new variable

```{r}
Bselect2[1:500, ]
summary(Bselect2$HSI)
```

#remove missing by removig rows if HSI value is N/A

```{r}
complete.cases(Bselect2$HSI)
```

```{r}
df_Bselect2<-Bselect2[complete.cases(Bselect2$HSI),]
```

#check if NA is removed

```{r}
df_Bselect2[1:500, ]
summary(df_Bselect2$HSI)
```

#NA is removed but there is an outlier as the max should only be 6, so removing the oulier. plotting a historgram and boxplot to see how many outliers there are

```{r}
ggplot(df_Bselect2)+
  aes(x=HSI)+ geom_histogram(bins=30L,fill="#0c4c8a")+
  theme_minimal()
```

```{r}
boxplot(df_Bselect2$HSI, ylab="HSI")
```

#identify the outlier

```{r}
boxplot.stats(df_Bselect2$HSI)$out
```

#Identify the rows correpding to the outlier

```{r}
out<-boxplot.stats(df_Bselect2$HSI)$out
out_ind<-which(df_Bselect2$HSI %in% c(out))
out_ind
```

#subsetting to remove rows with 10 from the df

```{r}
df_BSelect3<-subset(df_Bselect2, HSI!=10)
```

#check if outlier removed

```{r}
summary(df_BSelect3$HSI)
```

#catagorize HSI scores by 0 for scores 0-2, 1 for scores 3 to 6,
female(2) recoded to 0 keep male as is 1

```{r}
df_BSelect3$HSI[df_BSelect3$HSI>=0 & df_BSelect3$HSI<=2] <- 0

df_BSelect3$HSI[df_BSelect3$HSI>=3 & df_BSelect3$HSI<=6] <- 1

df_BSelect3$A01[df_BSelect3$A01>=2] <- 0
```

```{r}
summary(df_BSelect3)
```

#plot

```{r}
ggplot(df_BSelect3)+geom_bar(aes(x=A01, fill=HSI))
ggplot(df_BSelect3)+geom_bar(aes(x=A01, fill=HSI))+ggtitle(label="Heavy Smoking Index by Gender where 1 is Male and 0 is Female")
```

#regression y\~x

```{r}
fit1<-glm(HSI~A01,family=binomial(link="logit"),data=df_BSelect3)
```

```{r}
summary(fit1)
```

#The coeffecient of HSI is .11. The log of odds of being Male(1) is .11
times higher for those with medium addiction compared to those with low
addiction.

```{r}
print(fit1)
```

#save cleanted dataset as RDS

```{r}
saveRDS(df_BSelect3,file="df_BSelect3.RDS")
```

To cite other work (important everywhere, but likely happens first in
introduction), make sure your references are in the bibtex file
specified in the YAML header above (here
`dataanalysis_template_references.bib`) and have the right bibtex key.
Then you can include like this:

Examples of reproducible research projects can for instance be found in
[@mckay2020; @mckay2020a]

{{< pagebreak >}}

# Methods

*Describe your methods. That should describe the data, the cleaning
processes, and the analysis approaches. You might want to provide a
shorter description here and all the details in the supplement.*

## Data aquisition

*As applicable, explain where and how you got the data. If you directly
import the data from an online source, you can combine this section with
the next.*

## Data import and cleaning

*Write code that reads in the file and cleans it so it's ready for
analysis. Since this will be fairly long code for most datasets, it
might be a good idea to have it in one or several R scripts. If that is
the case, explain here briefly what kind of cleaning/processing you do,
and provide more details and well documented code somewhere (e.g. as
supplement in a paper). All materials, including files that contain
code, should be commented well so everyone can follow along.*

## Statistical analysis

*Explain anything related to your statistical analyses.*

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

*Use a combination of text/tables/figures to explore and describe your
data. Show the most important descriptive results here. Additional ones
should go in the supplement. Even more can be in the R and Quarto files
that are part of your project.*

Note the loading of the data providing a **relative** path using the
`../../` notation. (Two dots means a folder up). You never want to
specify an **absolute** path like `C:\ahandel\myproject\results\`
because if you share this with someone, it won't work for them since
they don't have that path. You can also use the `here` R package to
create paths. See examples of that below.

## Basic statistical analysis

*To get some further insight into your data, if reasonable you could
compute simple statistics (e.g. simple models with 1 predictor) to look
for associations between your outcome(s) and each individual predictor
variable. Though note that unless you pre-specified the outcome and main
exposure, any "p\<0.05 means statistical significance" interpretation is
not valid.*

## Full analysis

*Use one or several suitable statistical/machine learning methods to
analyze your data and to produce meaningful figures, tables, etc. This
might again be code that is best placed in one or several separate R
scripts that need to be well documented. You want the code to produce
figures and data ready for display as tables, and save those. Then you
load them here.*

getwd()

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

*Summarize what you did, what you found and what it means.*

## Strengths and Limitations

*Discuss what you perceive as strengths and limitations of your
analysis.*

## Conclusions

*What are the main take-home messages?*

*Include citations in your Rmd file using bibtex, the list of references
will automatically be placed at the end*

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers
published using a fully reproducible setup similar to the one shown in
this template.

Note that this cited reference will show up at the end of the document,
the reference formatting is determined by the CSL file specified in the
YAML header. Many more style files for almost any journal [are
available](https://www.zotero.org/styles). You also specify the location
of your bibtex reference file in the YAML. You can call your reference
file anything you like, I just used the generic word `references.bib`
but giving it a more descriptive name is probably better.

{{< pagebreak >}} library(pandoc)
