---
title: "Tidy Tuesday Exercise2"
output: 
  html_document:
    toc: FALSE
---

**To be filled..**

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(tidytuesdayR)#this installs the dataset
library(here)
library(knitr)
library(dplyr)
library(tidyr)
library(skimr)
library(haven)
library(memisc)
library(ggplot2)
library(reshape)
library(ggthemes)
library(DataExplorer)
library(table1)
```
**Download/access data**
```{r}
eggproduction  <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-04-11/egg-production.csv')
cagefreepercentages <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-04-11/cage-free-percentages.csv')
```

**Explore data**
#take a first look a the dataset
```{r}
view(cagefreepercentages)
view(eggproduction)
```
 ggplot2::ggplot(ggplot2::aes(x = observed_month, y = percent_hens)) +
    ggplot2::geom_point(color = color, size = 1) +
    ggplot2::geom_line(color = color) +
    ggplot2::scale_x_date(
      breaks = scales::date_breaks("1 year"),
      date_labels = "%Y"
#missing data
```{r}
plot_missing(cagefreepercentages)
```
# cagefree percentages seems like has almost half missing values 

```{r}
plot_missing(eggproduction)
```
#looks like the eggs production looks like a complete dataset with no missing values. 

```{r}
view(eggproduction)
```

#there seems to be an imbalance  where hatching eggs only has "all" observation, table eggs seem to have all three different process types. so removing the hatching eggs obsevation for a better analysis. 



#now lets understand what the missing values are about in the cage free percentages dataset 

```{r}
view(cagefreepercentages)
```
#It looks like those from Egg-Markets-Overview-2019-10-19.pdf dont have percent egg computed, since we dont have the denominator we cannt compute so best to remove NA's

```{r}
cagefreepercentages <- cagefreepercentages[complete.cases(cagefreepercentages), ]
view(cagefreepercentages)
```
#Eggfee percentages variabls include #data/percentegg/percenthen/source with n=54
#eggproduction variables include #producttype/productprocess/eggs/hens/sources with n=220
#by combining the two datasets we should have 

#lets plot some graphs and see if there are any relationships

```{r}
ggplot(cagefreepercentages,aes(observed_month))+ #basic graph object
  geom_line(aes(y=percent_hens), colour="red") + #layor 1
  geom_line(aes(y=percent_eggs), colour ="green")+ #layor 2
ggtitle("Percentages of Hens(RED) and eggs(Green), 2017-2021")+
 xlab("Year")+ #add x axis label
  ylab("Percentages")   #add y axis label

```
#seems like there is a relationship here with both hens and eggs percentages with an upward trend over time, note: I was not able to add the legend in the graph function 

#Now lets look at the eggproductions dataset 
```{r}
mosaicplot(prod_type~prod_process,data=eggproduction,col=c("Blue","Red","Pink")) 
```
#This figure illestrates the two different product types (hatching eggs and table eggs) and the three different process types caged-organic, caged-non organic and all. Hatching eggs contains only "all" value whereas table eggs has equal amount of caged organic, caged non-organic and all(combined). 
#still trying to find inspiration from this data. 
#Perhaps we can go ahead and merge these two datasets for a more complete picture. 

#Merge two datasets based on their common variable (inner join) observed month

```{r}
merged<- merge(x = cagefreepercentages, y = eggproduction, by = c("observed_month"))
```
#check your merged data

```{r}
view(merged)
```
#lets take a look 

```{r}
Eggdata <- (table1(~ n_hens + n_eggs  + factor(prod_process) +factor(prod_type)+percent_eggs + percent_hens , data=merged))

Eggdata <- t1kable(Eggdata)
Eggdata
```



```{r}
Eggdata <- (table1(~ factor(prod_process) + n_hens + n_eggs +percent_eggs + percent_hens  | prod_type, data=merged))

Eggdata <- t1kable(Eggdata)
Eggdata
```
The above table shows all prod process types are available for hatching eggs where as cage free organice and non organic are available only for table eggs.

Next steps , removing "all" observations from hatching eggs so that we are left with table eggs only with three prod process types. 
#here I got stuck with an error about deleting atomic verctors..

